### 2022-03-22 paper/output/document/thesis_3_implementation.tex
finished another read thrgh

\isamarkupsubsubsection{Logical Background%

@@

type, so the term $\neg M$ is not type correct. The expression $obligated \, \neg M$ merely provides 

@@

obligations, so I will add this as an axiom.%

@@

With this logical background, I can begin to implement the Formula of Universal Law, which, as defined 

@@

using the abbreviations above. While this attempt will not be consistent, I will use Isabelle's automatic

@@

contradicts this particular axiom offers insight into what the problem is. If the goal or action
or a maxim are equivalent to its circumstance, then prohibiting it is contradictory. If the maxim has already been 

@@

  Empty assignment \color{black}%
}%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
%
\endisadelimproof
%
\begin{isamarkuptext}%
My above investigation of \texttt{FUL0} shows that if the FUL holds for badly formed maxims, then 
it is inconsistent. This is not only a logical property of my system, but it also has philosophical
significance that coheres with Korsgaard's and O'Neill's interpretations of a maxim as a practical
guide to action \citep{actingforareason,actingonprinciple}. A maxim is a practical principle that guides how we behave in everyday life. A 
principle of the form ``When you are eating breakfast, eat breakfast in order to eat breakfast,'' is not 
practically relevant. No agent would ever need to act on such a principle. Morality helps agents decide
whether to act on a potential principle of action, but no agent would need to ask
``When I am eating breakfast, should I eat breakfast in order to eat breakfast?" It is the wrong kind of 
principle to be evaluating. It is not a well-formed maxim, so the categorical imperative cannot apply to it.

The fact that Isabelle revealed a philosophical insight about which kinds of maxims are well-formed
is an example of the power of computational tools to aid
philosophical progress. Nitpick and Sledgehammer helped me confirm that certain kinds
of circumstance, act, goal tuples are too badly formed for the categorical imperative to logically 
apply to them. The realization of this subtle problem would have been incredibly difficult without 
computational tools, and serves as evidence of the power of computational ethics. I discuss the philosophical
properties and implications of well-formed maxims and the power of computational ethics further in 
Section \ref{computationalethics}.

I complete my implementation by adding the consistent version of the FUL as an axiom.%
\end{isamarkuptext}\isamarkuptrue%
\isacommand{axiomatization}\isamarkupfalse%
\ \isakeyword{where}\ FUL{\isacharcolon}FUL%
\begin{isamarkuptext}%
This concludes my implementation of the Formula of Universal Law in Isabelle/HOL. My implementation
consistents of necessary logical background, first formalized in DDL and then implemented in Isabelle.
The code snippets in this chapter are a subset of over 100 lines of Isabelle/HOL code necessary to
complete this implementation. In Section \ref{testing} and Chapter \ref{applications}, I demonstrate how this
implementation can be tested and used to make moral judgements.%
\end{isamarkuptext}\isamarkuptrue%
%
\isadelimdocument
%
\endisadelimdocument
%
\isatagdocument
%
\isamarkupsubsection{Tests\label{testing}%
}
\isamarkuptrue%
%
\endisatagdocument
{\isafolddocument}%
%
\isadelimdocument
%
\endisadelimdocument
%
\begin{isamarkuptext}%
In addition to an implementation of automated Kantian ethics, I also contribute a testing framework 
to evaluate how well my implementation coheres with philosophical literature. This testing architecture 
makes the notion of ``philosophical faithfulness'' precise. Each test consists of a sentence in my logic 
and an expected outcome, where the possible outcomes are proving or refuting the sentence. For example, 
one such sentence is that obligations cannot contradict each other. To run the tests, I attempt to
 prove or refute each test sentence in my logic.  Because these tests are derived from moral 
intuition and philosophical literature, they evaluate how well my system reflects philosophical
literature. Running the tests on my implementation consisted of approximately 400 lines of Isabelle code.

The test framework can be expanded by adding more test sentences and can guide 
implementations of other parts of Kantian ethics or other ethical theories. As I was implementing my 
formalization, I checked it against the testing framework, performing test-driven development for automated 
ethics.

I use my testing framework to show that my formalization and implementation of Kantian ethics outperforms 

@@

on top of which I build other implementations. DDL can express obligation, but has no knowledge of any 

@@

of the FUL and the second formulation
of the categorical imperative \citep{kroy}. His formalization is based on Hintikka's deontic logic, which is a different,

@@

\noindent \textbf{FUL Stronger than DDL} The FUL should not hold in raw DDL, which I use a control group. 

@@

my formalization satisfy this property, but raw DDL does not. Below I run
this test for my formalization.%

@@

\noindent \textbf{Obligations Never Contradict} Contradictory obligations

@@

Kant only briefly discusses contradictory obligations in \emph{Metaphysics of Morals}, where he argues that 

@@

Kroy's formalization allow contradictory obligations, but I explicitly add an axiom to my formalization

@@

\noindent \textbf{Distributive Property for Obligations Holds} Another property related to contradictory obligations is the distributive property for the obligation

@@

another property that we expect to hold. The rough English translation of  $O \{ A \wedge B \} $ is ``you are obligated to 

@@

and you are obligated to do B.'' We think those English sentences mean the same thing, and they should mean 

@@

@@

passes for my interpretation because I require that obligations don't contradict as an axiom.%

@@

@@

\noindent \textbf{Prohibits Actions That Are Impossible to Universalize} Recall that 
the logical contradiction interpretation of the Formula of Universal Law prohibits lying because in a world 
where everyone simultaneously lies, lying is impossible. In other words, not everyone can simultaneously
lie because the institution of lying and believing would break down. In Section \ref{praccon}, 
I recreated Korsgaard's argument for why the logical contradiction interpretation is weaker than what the
Formula of Universal Law should actually require. Therefore, any implementation of the FUL should be 
able to show that the actions prohibited by the logical contradiction interpretation are prohibited, 
because the set of actions prohibited by the practical contradiction interpretation is a superset of these.
The FUL should show that actions that cannot possibly be universalized are prohibited, because those acts cannot be willed in 
a world where they are universalized. This property fails to hold in both raw DDL
and Kroy's formalization, but holds for my formalization. Showing that this property holds
for my formalization required significant logical work and the full code is presented in Appendix \ref{weirdtests}.%
\end{isamarkuptext}\isamarkuptrue%
%
\begin{isamarkuptext}%
\noindent \textbf{Robust Representation of Maxims} Kant does not evaluate the correctness of acts, but rather of maxims. Therefore, any 
faithful formalization of the categorical imperative must evaluate maxims, not acts. This requires 
representing a maxim and making it the input to the obligation operator, which only my implementation 
does. Because my implementation includes the notion of a maxim, it is able to perform sophisticated 
reasoning as demonstrated in Sections \ref{joking} and \ref{murderer}. Staying faithful to the philosophical 
literature enables my system to make more reliable judgements.

\medskip%
\end{isamarkuptext}\isamarkuptrue%
%
\begin{isamarkuptext}%
\noindent \textbf{Can Prohibit Conventional and Natural Acts} When arguing for the practical contradiction interpretation,
Korsgaard makes a distinction between conventional and natural acts \citep{KorsgaardFUL}. 
A conventional act relies on a convention, like the 
convention that a promise is a commitment, whereas a natural act is possible simply because of the laws 
of the natural world. Conventional acts exist within a practice, which is ``comprised of certain rules, 
and its existence (where it is not embodied in an institution with sanctions) consists in the general 
acknowledgement and following of those rules'' \cite[10]{KorsgaardFUL}. Promising is a conventional act 
because it exists as a practice. Murder, on the other hand, is an example of a natural act because 
its existence only depends on the laws of nature\cite[11]{KorsgaardFUL}.

@@

Both raw DDL and Kroy's interpretation fail to show the wrongness of conventional or natural acts. 

@@

this test in Appendix \ref{weirdtests}.%

@@

%:%515=360%:%
%:%516=360%:%

@@

%:%519=362%:%

@@

%:%523=365%:%
%:%525=367%:%
%:%526=367%:%

@@

%:%529=369%:%

@@

%:%531=370%:%

@@

%:%534=372%:%

@@

%:%536=373%:%
%:%538=373%:%
%:%542=373%:%
%:%543=373%:%
%:%545=374%:%

@@

%:%561=382%:%

@@

@@

%:%583=403%:%

@@

%:%586=406%:%

@@

%:%598=411%:%
%:%610=413%:%
%:%611=414%:%

@@

%:%637=441%:%
%:%638=442%:%

@@

%:%645=448%:%
%:%646=449%:%

@@

%:%658=459%:%
%:%659=460%:%

@@

@@

%:%674=474%:%
%:%675=475%:%
%:%678=476%:%
%:%682=476%:%
%:%683=476%:%
%:%685=477%:%
%:%686=478%:%

@@

%:%688=479%:%
%:%690=480%:%
%:%691=481%:%

@@

%:%699=483%:%
%:%702=484%:%
%:%703=485%:%
%:%704=485%:%
%:%705=486%:%
%:%708=487%:%
%:%712=487%:%
%:%713=487%:%
%:%715=488%:%
%:%716=489%:%
%:%726=491%:%
%:%727=492%:%

@@

@@

%:%749=513%:%
%:%750=514%:%
%:%753=515%:%
%:%757=515%:%
%:%758=515%:%
%:%759=515%:%
%:%761=516%:%
%:%767=516%:%
%:%770=517%:%
%:%771=518%:%
%:%772=518%:%
%:%773=519%:%
%:%774=520%:%
%:%777=521%:%
%:%781=521%:%
%:%782=521%:%
%:%783=521%:%
%:%785=522%:%
%:%786=523%:%
%:%787=524%:%
%:%797=526%:%
%:%798=527%:%
%:%799=528%:%

@@

%:%806=536%:%

@@

%:%810=537%:%
%:%814=537%:%
%:%815=537%:%
%:%817=538%:%
%:%818=539%:%
%:%819=539%:%
%:%821=540%:%
%:%831=542%:%
%:%832=543%:%
%:%833=544%:%

@@

@@

%:%840=550%:%
%:%841=551%:%
%:%842=552%:%
%:%845=553%:%
%:%849=553%:%
%:%850=553%:%
%:%851=553%:%
%:%853=554%:%
%:%854=555%:%
%:%855=556%:%
%:%865=558%:%
%:%866=559%:%
%:%868=561%:%
%:%869=561%:%
%:%870=562%:%
%:%871=563%:%
%:%874=564%:%
%:%878=564%:%
%:%879=564%:%
%:%880=564%:%
%:%882=565%:%
%:%892=567%:%
%:%893=568%:%
%:%894=569%:%
%:%895=570%:%
%:%896=571%:%
%:%897=572%:%
%:%898=573%:%
%:%899=574%:%
%:%900=575%:%
%:%901=576%:%
%:%902=577%:%
%:%903=578%:%
%:%904=579%:%
%:%905=580%:%
%:%909=584%:%
%:%910=585%:%
%:%911=586%:%
%:%912=587%:%
%:%913=588%:%
%:%914=589%:%
%:%915=590%:%
%:%916=591%:%
%:%920=594%:%
%:%921=595%:%
%:%922=596%:%
%:%923=597%:%
%:%924=598%:%
%:%925=599%:%
%:%926=600%:%
%:%927=601%:%
%:%928=602%:%
%:%929=603%:%
%:%930=604%:%
%:%931=605%:%
%:%932=606%:%
%:%933=607%:%
%:%934=608%:%
%:%935=609%:%
%:%936=610%:%
%:%937=611%:%
%:%938=612%:%
%:%939=613%:%
%:%940=614%:%
%:%941=615%:%
%:%942=616%:%
%:%943=617%:%
%:%944=618%:%

>>>>

\isamarkupsubsubsection{Subjects and Acts%

@@

type, so the term $\neg M$ is not type correct. The expression $\text{obligated} \, \neg M$ merely provides 

@@

obligations, so I will add an axiom disallowing contradictory obligations.%

@@

With this logical background, I can implement the Formula of Universal Law, which, as defined 

@@

using the abbreviations above. While this attempt is not consistent, I use Isabelle's automatic

@@

contradicts this particular axiom offers insight into what the problem is: the FUL must be obligating 
a maxim in which the act or goal contradicts the circumstances. Precisely, if the goal or action
or a maxim are equivalent to its circumstances, then prohibiting it is contradictory. If the maxim has already been 

@@

  Empty assignment \color{black}%
}%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
%
\endisadelimproof
%
\begin{isamarkuptext}%
My above investigation of \texttt{FUL0} shows that if the FUL holds for badly formed maxims, then 
it is inconsistent. This is not only a logical property of my system, but it also has philosophical
significance that coheres with Korsgaard's and O'Neill's interpretations of a maxim as a practical
guide to action \citep{actingforareason,actingonprinciple}. A maxim is a practical principle that guides how we behave in everyday life. A 
principle of the form ``When you are eating breakfast, eat breakfast in order to eat breakfast,'' is not 
practically relevant. Morality helps agents decide
whether to act on a potential principle of action, but no agent would need to ask
``When I am eating breakfast, should I eat breakfast in order to eat breakfast?" Because it is not a 
well-formed maxim, it is the wrong kind of principle to be evaluating, so the categorical imperative 
cannot apply to it.

The fact that Isabelle revealed a philosophical insight about which kinds of maxims are well-formed
is an example of the power of computational tools to aid
philosophical progress. Nitpick and Sledgehammer helped me confirm that certain kinds
of circumstance, act, goal tuples are too badly formed for the categorical imperative to logically 
apply to them. The realization of this subtle problem would have been incredibly difficult without 
computational tools, and serves as evidence of the power of computational ethics. I further discuss the philosophical
properties and implications of well-formed maxims and the power of computational ethics in 
Section \ref{computationalethics}.

I complete my implementation by adding the consistent version of the FUL as an axiom.%
\end{isamarkuptext}\isamarkuptrue%
\isacommand{axiomatization}\isamarkupfalse%
\ \isakeyword{where}\ FUL{\isacharcolon}FUL%
\begin{isamarkuptext}%
This concludes my implementation of the Formula of Universal Law in Isabelle/HOL. My implementation
consistents of necessary logical background, first formalized in DDL and then implemented in Isabelle.
The code snippets in this chapter are a subset of over 100 lines of Isabelle/HOL code necessary to
complete this implementation. In Section \ref{testing} and Chapter \ref{applications}, I demonstrate how this
implementation can be tested and used to make moral judgements.%
\end{isamarkuptext}\isamarkuptrue%
%
\isadelimdocument
%
\endisadelimdocument
%
\isatagdocument
%
\isamarkupsubsection{Tests\label{testing}%
}
\isamarkuptrue%
%
\endisatagdocument
{\isafolddocument}%
%
\isadelimdocument
%
\endisadelimdocument
%
\begin{isamarkuptext}%
In addition to an implementation of automated Kantian ethics, I also contribute a testing framework 
to evaluate how well my implementation coheres with philosophical literature. This testing architecture 
makes the notion of ``philosophical faithfulness'' precise. Each test consists of a sentence in my logic\footnote{Some
tests also require additional logical background, explained below.} and an expected outcome, where 
the possible outcomes are proving or refuting the sentence. For example, 
one such sentence is that obligations cannot contradict each other. To run the tests, I attempt to
 prove or refute each test sentence in my logic.  Because these tests are derived from moral 
intuition and philosophical literature, they evaluate how well my system reflects philosophical
literature. Running the tests on my implementation consisted of approximately 400 lines of Isabelle code.

The testing framework can be expanded by adding more test sentences and can guide 
implementations of other parts of Kantian ethics or other ethical theories. As I was implementing my 
formalization, I checked it against the testing framework, performing test-driven development for automated 
ethics.

I use my testing framework to show that my formalization and implementation of Kantian ethics outperform 

@@

on top of which I build other implementations. DDL can express obligation, but does not include any 

@@

of the FUL \citep{kroy}. His formalization is based on Hintikka's deontic logic, which is a different,

@@

\noindent \textbf{FUL Stronger than DDL} The FUL should not hold in raw DDL, which I use as a control group. 

@@

my implementation satisfy this property, but raw DDL does not. Below I run
this test for my implementation.%

@@

\noindent \textbf{Obligations Never Contradict} There are two reasons that Kantian ethics cannot
require contradictory obligations. First, contradictory obligations

@@

Second, Kant briefly discusses contradictory obligations in \emph{Metaphysics of Morals}, where he argues that 

@@

Kroy's formalization allow contradictory obligations, but I explicitly add an axiom to my implementation

@@

\noindent \textbf{Distributive Property for Obligations Holds} Another property related to 
contradictory obligations is the distributive property for the obligation operator over the and 

@@

another property that should hold. The rough English translation of  $O \{ A \wedge B \}$ is ``you are obligated to 

@@

and you are obligated to do B.'' Both those English sentences mean the same thing, and they should mean 

@@

\isamarkupcmt{This lemma is trivially true by the added axiom, but notice that this proof does not use that axiom.
This proof uses axiom O\_diamond and the assumptions and thus also holds in DDL itself, even without 
an axiom prohibiting contradictory obligations.%

@@

passes for my implementation because I require that obligations don't contradict as an axiom.%

@@

\isamarkupcmt{This proof and the above lemma together imply that the distributive property holds if and only if 
there are no contradictory obligations.%

@@

\noindent \textbf{Prohibits Actions That Are Impossible to Universalize} Recall that 
the logical contradiction interpretation of the Formula of Universal Law prohibits lying because in a world 
where everyone simultaneously lies, lying is impossible. In other words, not everyone can simultaneously
lie because the institution of lying and believing would break down. In Section \ref{praccon}, 
I recreated Korsgaard's argument for why the logical contradiction interpretation is weaker than what the
Formula of Universal Law should actually require. Therefore, any implementation of the FUL should be 
able to show that the actions prohibited by the logical contradiction interpretation are prohibited, 
because the set of actions prohibited by the practical contradiction interpretation is a superset of these.
The FUL should show that actions that cannot possibly be universalized are prohibited, because those acts cannot be willed in 
a world where they are universalized. This property fails to hold in both raw DDL
and Kroy's formalization, but holds for my implementation. Showing that this property holds
for my formalization required additional logical background and the full code is presented in Appendix \ref{weirdtests}.%
\end{isamarkuptext}\isamarkuptrue%
%
\begin{isamarkuptext}%
\noindent \textbf{Robust Representation of Maxims} Kant does not evaluate the correctness of acts, but rather of maxims. Therefore, any 
faithful formalization of the categorical imperative must evaluate maxims, not acts. This requires 
representing a maxim and making it the input to the obligation operator, which only my implementation 
does. Because my implementation includes the notion of a maxim, it is able to perform sophisticated 
reasoning as demonstrated in Sections \ref{joking} and \ref{murderer}. Staying faithful to the philosophical 
literature enables my system to make more reliable judgements.

\medskip%
\end{isamarkuptext}\isamarkuptrue%
%
\begin{isamarkuptext}%
\noindent \textbf{Can Prohibit Conventional and Natural Acts} When arguing for the practical contradiction interpretation,
Korsgaard makes a distinction between conventional and natural acts \citep{KorsgaardFUL}. 
A conventional act relies on a convention, like the 
convention that a promise is a commitment, whereas a natural act is possible simply because of the laws 
of the natural world. Conventional acts exist within a practice, which is ``comprised of certain rules, 
and its existence (where it is not embodied in an institution with sanctions) consists in the general 
acknowledgement and following of those rules'' \cite[10]{KorsgaardFUL}. Promising is a conventional act 
because it exists as a practice. Murder, on the other hand, is an example of a natural act because 
its existence only depends on the laws of nature \cite[11]{KorsgaardFUL}.

@@

Both raw DDL and Kroy's formalization fail to show the wrongness of conventional or natural acts. 

@@

this test in Appendix \ref{weirdtests}. Because I formalize the philosophically-accepted version 
of the FUL, my implementation passes this test.%

@@

%:%514=359%:%
%:%516=361%:%

@@

%:%518=362%:%

@@

%:%521=364%:%
%:%524=366%:%
%:%526=368%:%

@@

%:%528=369%:%

@@

%:%531=371%:%

@@

%:%534=373%:%

@@

%:%536=374%:%
%:%537=374%:%
%:%539=374%:%
%:%543=374%:%
%:%544=374%:%

@@

%:%552=381%:%

@@

%:%581=402%:%

@@

%:%584=405%:%

@@

@@

%:%590=410%:%
%:%591=411%:%
%:%600=413%:%

@@

%:%635=438%:%
%:%636=439%:%

@@

%:%643=447%:%
%:%644=448%:%

@@

%:%655=458%:%
%:%656=459%:%

@@

%:%672=473%:%

@@

%:%675=476%:%
%:%676=476%:%
%:%677=477%:%
%:%680=478%:%
%:%684=478%:%
%:%685=478%:%

@@

%:%688=480%:%
%:%689=481%:%
%:%690=481%:%

@@

%:%694=484%:%
%:%695=485%:%
%:%701=485%:%
%:%704=486%:%
%:%705=487%:%
%:%706=487%:%
%:%707=488%:%
%:%710=489%:%
%:%714=489%:%
%:%715=489%:%
%:%717=490%:%
%:%718=491%:%

@@

%:%747=512%:%

@@

%:%749=514%:%
%:%751=516%:%
%:%752=516%:%
%:%753=517%:%
%:%756=518%:%
%:%760=518%:%
%:%761=518%:%
%:%762=518%:%
%:%764=519%:%
%:%770=519%:%
%:%773=520%:%
%:%774=521%:%
%:%775=521%:%
%:%776=522%:%
%:%777=523%:%
%:%780=524%:%
%:%784=524%:%
%:%785=524%:%
%:%786=524%:%
%:%788=525%:%
%:%789=526%:%
%:%790=527%:%

@@

%:%805=534%:%
%:%806=535%:%

@@

%:%809=539%:%
%:%810=539%:%
%:%813=540%:%
%:%817=540%:%
%:%818=540%:%
%:%820=541%:%
%:%821=542%:%
%:%822=542%:%
%:%824=543%:%

@@

%:%838=549%:%

@@

%:%840=551%:%
%:%841=552%:%
%:%843=554%:%
%:%844=554%:%
%:%845=555%:%
%:%846=556%:%
%:%849=557%:%
%:%853=557%:%
%:%854=557%:%
%:%855=557%:%
%:%857=558%:%
%:%858=559%:%
%:%859=560%:%
%:%860=560%:%
%:%862=561%:%
%:%863=562%:%
%:%864=563%:%
%:%874=565%:%
%:%875=566%:%
%:%877=568%:%
%:%878=568%:%
%:%879=569%:%
%:%880=570%:%
%:%883=571%:%
%:%887=571%:%
%:%888=571%:%
%:%889=571%:%
%:%891=572%:%
%:%892=572%:%
%:%894=573%:%
%:%895=574%:%
%:%905=576%:%
%:%906=577%:%
%:%907=578%:%
%:%908=579%:%
%:%909=580%:%
%:%910=581%:%
%:%911=582%:%
%:%912=583%:%
%:%913=584%:%
%:%914=585%:%
%:%915=586%:%
%:%916=587%:%
%:%917=588%:%
%:%918=589%:%
%:%922=593%:%
%:%923=594%:%
%:%924=595%:%
%:%925=596%:%
%:%926=597%:%
%:%927=598%:%
%:%928=599%:%
%:%929=600%:%
%:%933=603%:%
%:%934=604%:%
%:%935=605%:%
%:%936=606%:%
%:%937=607%:%
%:%938=608%:%
%:%939=609%:%
%:%940=610%:%
%:%941=611%:%
%:%942=612%:%
%:%943=613%:%
%:%944=614%:%
%:%945=615%:%
%:%946=616%:%
%:%947=617%:%
%:%948=618%:%
%:%949=619%:%
%:%950=620%:%
%:%951=621%:%
%:%952=622%:%
%:%953=623%:%
%:%954=624%:%
%:%955=625%:%
%:%956=626%:%
%:%957=627%:%
%:%958=628%:%
